### 2022-08-22 doc/TrashData.tex
Gloria's 8/12 edits.

\begin{table}

\caption{\label{tab:trashdata}Republic/Santek Merger Simulation Inputs. Prices and margins are reported in dollars, while volume is reported in thousands of pounds.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rrcccccr}
\toprule
Disposal Firm & Collection Firm & Volume & Disposal Price & Disposal Margin & Collection Margin & Collection Price & Volume (000s)\\
\midrule
Republic & Republic & 164728 & 42 & 0 & 44 & 179 & 165\\
\cmidrule{1-8}
 & Republic & 34008 & 36 & 20 &  & 173 & 34\\

 & Santek & 217727 & 16 & 0 &  & 134 & 218\\

 & WM-ADS & 30492 & 36 & 20 &  & 221 & 30\\

\multirow{-4}{*}{\raggedleft\arraybackslash Santek} & Regional & 30492 & 36 & 20 &  & 221 & 30\\
\cmidrule{1-8}
 & Republic & 19289 & 25 & 14 &  & 162 & 19\\

 & WasteConn & 47786 & 11 & 0 &  & 110 & 48\\

 & WM-ADS & 17294 & 25 & 14 &  & 210 & 17\\

\multirow{-4}{*}{\raggedleft\arraybackslash WasteConn} & Regional & 17294 & 25 & 14 &  & 210 & 17\\
\bottomrule
\end{tabular}}
\end{table}