### 2022-08-22 doc/TrashData.tex
minor edits to trashsims table

\begin{table}

\caption{\label{tab:trashdata}Republic/Santek Merger Simulation Inputs. Prices and margins are reported in dollars, while volume is reported in thousands of pounds.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rrccccc}
\toprule
Disposal Firm & Collection Firm & Volume & Disposal Price & Disposal Margin & Collection Margin & Collection Price\\
\midrule
Republic & Republic & 165 & 42 & 0 & 44 & 179\\
\cmidrule{1-7}
 & Republic & 34 & 36 & 20 &  & 173\\

 & Santek & 218 & 16 & 0 &  & 134\\

 & WM-ADS & 30 & 36 & 20 &  & 221\\

\multirow{-4}{*}{\raggedleft\arraybackslash Santek} & Regional & 30 & 36 & 20 &  & 221\\
\cmidrule{1-7}
 & Republic & 19 & 25 & 14 &  & 162\\

 & WasteConn & 48 & 11 & 0 &  & 110\\

 & WM-ADS & 17 & 25 & 14 &  & 210\\

\multirow{-4}{*}{\raggedleft\arraybackslash WasteConn} & Regional & 17 & 25 & 14 &  & 210\\
\bottomrule
\end{tabular}}
\end{table}