### 2024-09-30 doc/TrashData.tex
replaced downstream prices with costs in output table.

\caption{\label{tab:trashdata}Republic/Santek Merger Simulation Inputs. Volume is reported in thousands of pounds, while prices and margins are reported in dollars.}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{rrccccc}
\toprule
Disposal Firm & Collection Firm & Volume & Disposal Price & Disposal Margin & Collection Margin & Collection Cost\\
\midrule
Republic & Republic & 165 & 42 & 0 & 44 & 93\\
\cmidrule{1-7}
 & Republic & 34 & 36 & 20 &  & 93\\

 & Santek & 218 & 16 & 0 &  & 74\\

 & WMADS & 30 & 36 & 20 &  & 148\\

\multirow{-4}{*}{\raggedleft\arraybackslash Santek} & Regional & 30 & 36 & 20 &  & 148\\
\cmidrule{1-7}
 & Republic & 19 & 25 & 14 &  & 93\\

 & WasteConn & 48 & 11 & 0 &  & 63\\

 & WMADS & 17 & 25 & 14 &  & 148\\

\multirow{-4}{*}{\raggedleft\arraybackslash WasteConn} & Regional & 17 & 25 & 14 &  & 148\\
\bottomrule
\end{tabular}}